Efficiently summarizing relationships in large samples: a general duality between statistics of genealogies and genomes

### 2019-09-19 treestats_paper.tex
.

% visualization and time decomposition adding a new dimension
Genomic data are naturally distributed on two axes: along the genome, and across geography.
The tree sequence extends this to a third axis: time.
A great many methods in population genetics aim to describe aspects of history,
and accurate (or at least unbiased) tree sequence inference
may shift the focus of these problems from inference to descriptive analysis.
The methods developed here distribute the contribution to various statistics across each tree,
and so could also be used to summarize contributions to various statistics across time.
This could provide, for instance, the time distribution of mutations or branches
contributing positively or negatively to $f_4$ statistics of introgression,
enabling historical interpretations of these signals.
% This additional mode of reporting statistics is an important subject for future work.
% moving past a few summaries in discrete groups to a genome/time/space view of relatedness
The computational toolbox of population genetics
is still mostly composed of statistics orignally designed for analysis of a handful of loci
in a small number of discrete, mostly separated populations.
Both our data and our understanding of the world have moved beyond this.
We hope that the tools developed here will help make it possible
to visualize and analyze genetic variation and genealogical relatedness
along the genome, across space, and through time.

>>>>

% visualization and time decomposition adding a new dimension
Genomic data are naturally distributed on two axes: along the genome, and across geography.
The tree sequence extends this to a third axis: time.
A great many methods in population genetics aim to describe aspects of history,
and accurate (or at least unbiased) tree sequence inference
may shift the focus of these problems from inference to descriptive analysis.
The methods developed here distribute the contribution to various statistics across each tree,
and so could also be used to summarize contributions to various statistics across time.
This could provide, for instance, the time distribution of mutations or branches
contributing positively or negatively to $f_4$ statistics of introgression,
enabling historical interpretations of these signals.
% This additional mode of reporting statistics is an important subject for future work.
% moving past a few summaries in discrete groups to a genome/time/space view of relatedness
The computational toolbox of population genetics
is still mostly composed of statistics orginally designed for analysis of a handful of loci
in a small number of discrete, mostly separated populations.
Both our data and our understanding of the world have moved beyond this.
We hope that the tools developed here will help make it possible
to visualize and analyze genetic variation and genealogical relatedness
along the genome, across space, and through time.
